---
title: "IntroductiontoDataSience"
author: "Michael"
date: "1 5 2021"
output: pdf_document
---


---
title: "IntroToDataScience"
author: "Michael"
date: "26 3 2021"
output: pdf_document
---
#Load data +clean

##tweets data august
```{r}
#rm(list = ls())
library(tidyverse)

## load data ####august
covid19_tweets <- read.csv("covid19_tweets.csv", comment.char = "#")
glimpse(covid19_tweets)
#covid19_tweets$date <- as.Date(covid19_tweets$date)
#covid19_tweets$is_retweet <- as.logical(covid19_tweets$is_retweet)
#covid19_tweets$user_friends <- as.numeric(covid19_tweets$user_friends)
#covid19_tweets$user_followers <- as.numeric(covid19_tweets$user_followers)
#covid19_tweets$user_favourites <- as.numeric(covid19_tweets$user_favourites)
#covid19_tweets$user_verified <- as.logical(covid19_tweets$user_verified)

head(covid19_tweets)
summary(covid19_tweets)
```


##tweets data april
```{r}
covid_tweets_april <- data.frame()

filePaths <- list.files("C:/Users/micha/Desktop/Semester8/Introduction to Data Sience/Project_Group/Neuer Ordner/Data/covid19_tweets_april",  full.names = TRUE)
covid_tweets_april <- do.call(rbind, lapply(filePaths, read.csv))

#View(covid_tweets_april)
```


##make sure dates are safed as dates remove NA's
```{r}
covid19_tweets$date <- as.Date(covid19_tweets$date, format = "%Y-%m-%d")
covid_tweets_april$created_at <- as.Date(covid_tweets_april$created_at, format = "%Y-%m-%d")

covid19_tweets_august <-data.frame()
covid19_tweets_april <- data.frame()

covid19_tweets_august <- covid19_tweets[complete.cases(covid19_tweets$date),]
covid19_tweets_april <- covid_tweets_april[complete.cases(covid_tweets_april$created_at),]
```


#remove not used colums: using colums(date, text, country, ??any thing else??)
```{r}
new_covid19_tweets_august <- data.frame()
new_covid19_tweets_april <- data.frame()

new_covid19_tweets_april <- subset(covid19_tweets_april, select = -c(source:retweet_count, followers_count:verified))
new_covid19_tweets_august <- subset(covid19_tweets_august, select = -c(user_created:user_verified,user_description, hashtags, source:is_retweet))

library(dplyr)
new_covid19_tweets_april <- new_covid19_tweets_april %>% rename(user_name = screen_name, date = created_at , text = text)

```

##combine location colums in april data
```{r}

paste_noNA <- function(user_location,sep=", ") {
gsub(", " ,sep, toString(user_location[!is.na(user_location) & user_location!="" & user_location!="NA"] ) ) }

sep=" "
new_covid19_tweets_april$user_location <- apply( new_covid19_tweets_april[ ,c(6:8) ] , 1 , paste_noNA , sep=sep)
View(new_covid19_tweets_april)

new_covid19_tweets_april <- subset(new_covid19_tweets_april, select = -c(status_id, user_id, country_code, place_full_name, place_type, lang))

```


## select only Germany and United Kingdom ####

```{r}
## select only Germany and United Kingdom ## august ##
# using regex and str_detect from stringr

library(tidyverse)

# cleaning for regex: space padding for exact match
new_covid19_tweets_august$user_location <- paste(" ", new_covid19_tweets_august$user_location, " ")


tweetsUK_august <- filter(
    new_covid19_tweets_august,
    str_detect(user_location, regex(" United Kingdom ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Britain ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Wales ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Northern Ireland ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Scotland ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" England ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" UK ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" GB ", ignore_case = TRUE)) 
)

##Look if there are other labeling for the countries??

tweetsDE_august <- filter(
    new_covid19_tweets_august,
    str_detect(user_location, regex(" Germany ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Deutschland ", ignore_case = TRUE))
)

tweetsCH_august <- filter(
    new_covid19_tweets_august,
    str_detect(user_location, regex(" Switzerland ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" CH ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Svizzera ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Swiss ", ignore_case = TRUE))
)

tweetsIT_august <- filter(
    new_covid19_tweets_august,
    str_detect(user_location, regex(" Italy ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Italia ", ignore_case = TRUE)) 
)

## clean the tweet location ####
tweetsDE_august$user_location <- rep("DE", nrow(tweetsDE_august))
tweetsUK_august$user_location <- rep("UK", nrow(tweetsUK_august))
tweetsCH_august$user_location <- rep("CH", nrow(tweetsCH_august))
tweetsIT_august$user_location <- rep("IT", nrow(tweetsIT_august))



## filtered set ####
filtered_tweets_august <- rbind(tweetsUK_august, tweetsDE_august, tweetsCH_august, tweetsIT_august)
filtered_tweets_august$user_location <- as.factor(filtered_tweets_august$user_location)
```


```{r}
## select only Germany and United Kingdom ## april ##
# using regex and str_detect from stringr

library(tidyverse)

# cleaning for regex: space padding for exact match
new_covid19_tweets_april$user_location <- paste(" ", new_covid19_tweets_april$user_location, " ")


tweetsUK_april <- filter(
    new_covid19_tweets_april,
    str_detect(user_location, regex(" United Kingdom ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Britain ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Wales ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Northern Ireland ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Scotland ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" England ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" UK ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" GB ", ignore_case = TRUE)) 
)

##Look if there are other labeling for the countries??

tweetsDE_april <- filter(
    new_covid19_tweets_april,
    str_detect(user_location, regex(" Germany ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Deutschland ", ignore_case = TRUE))
)

tweetsCH_april <- filter(
    new_covid19_tweets_april,
    str_detect(user_location, regex(" Switzerland ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" CH ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Svizzera ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Swiss ", ignore_case = TRUE))
)

tweetsIT_april <- filter(
    new_covid19_tweets_april,
    str_detect(user_location, regex(" Italy ", ignore_case = TRUE)) |
        str_detect(user_location, regex(" Italia ", ignore_case = TRUE)) 
)

## clean the tweet location ####
tweetsDE_april$user_location <- rep("DE", nrow(tweetsDE_april))
tweetsUK_april$user_location <- rep("UK", nrow(tweetsUK_april))
tweetsCH_april$user_location <- rep("CH", nrow(tweetsCH_april))
tweetsIT_april$user_location <- rep("IT", nrow(tweetsIT_april))

## filtered set ####
filtered_tweets_april <- rbind(tweetsUK_april, tweetsDE_april, tweetsCH_april, tweetsIT_april)
filtered_tweets_april$user_location <- as.factor(filtered_tweets_april$user_location)
```

```{r}
filtered_tweets <-full_join(filtered_tweets_april, filtered_tweets_august)

```

#exporting Cleaned data
```{r}

write.csv(filtered_tweets, file = "filtered_tweets")
write.csv(filtered_tweets_april, file = "filtered_tweets_april")
write.csv(filtered_tweets_august, file = "filtered_tweets_august")

```

#############################################################################################################
#############################################################################################################

#mitigation measurements
##Load Data
```{r}
###https://www.kaggle.com/barun2104/government-measures-to-combat-covid19### --> gov

acaps_covid19_government_measures_dataset <- read.csv("C:/Users/micha/Desktop/Semester8/Introduction to Data Sience/Project_Group/Neuer Ordner/Data/government-measures-to-combat-covid19/acaps_covid19_government_measures_dataset.csv")

```

##filtering by country UK + De + CH + IT

###https://www.kaggle.com/barun2104/government-measures-to-combat-covid19### --> gov
```{r}
## select only Germany and United Kingdom ## 
# using regex and str_detect from stringr

library(tidyverse)

# cleaning for regex: space padding for exact match
acaps_covid19_government_measures_dataset$COUNTRY <- paste(" ", acaps_covid19_government_measures_dataset$COUNTRY, " ")


mitigationUK_GOV <- filter(
    acaps_covid19_government_measures_dataset,
    str_detect(COUNTRY, regex(" United Kingdom ", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex(" Britain ", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex(" Wales ", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex(" Northern Ireland ", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex(" Scotland ", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex(" England ", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex(" UK ", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex(" GB ", ignore_case = TRUE)) 
)

##Look if there are other labeling for the countries??

mitigationDE_GOV <- filter(
    acaps_covid19_government_measures_dataset,
    str_detect(COUNTRY, regex("Germany", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex("Deutschland", ignore_case = TRUE))|
        str_detect(COUNTRY, regex("DE", ignore_case = TRUE))
)

mitigationCH_GOV <- filter(
    acaps_covid19_government_measures_dataset,
    str_detect(COUNTRY, regex(" Switzerland ", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex(" CH ", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex(" Svizzera ", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex(" Swiss ", ignore_case = TRUE))
)

mitigationIT_GOV <- filter(
    acaps_covid19_government_measures_dataset,
    str_detect(COUNTRY, regex(" Italy ", ignore_case = TRUE)) |
        str_detect(COUNTRY, regex(" Italia ", ignore_case = TRUE)) 
)

## clean the tweet location ####
mitigationDE_GOV$COUNTRY <- rep("DE", nrow(mitigationDE_GOV))
mitigationUK_GOV$COUNTRY <- rep("UK", nrow(mitigationUK_GOV))
mitigationCH_GOV$COUNTRY <- rep("CH", nrow(mitigationCH_GOV))
mitigationIT_GOV$COUNTRY <- rep("IT", nrow(mitigationIT_GOV))

## filtered set ####
filtered_Mitigation_gov <- rbind(mitigationUK_GOV, mitigationDE_GOV, mitigationCH_GOV, mitigationIT_GOV)
filtered_Mitigation_gov$COUNTRY <- as.factor(filtered_Mitigation_gov$COUNTRY)
```


##remove not used colums
    ```{r}
#filtered_Mitigation_con1 <- subset(filtered_Mitigation_con, select = -c(ï..ID, Applies.To, Exceptions, Implementing.City, Implementing.State.Province, Quantity, Source, Target.city, Target.country, Target.region, Target.state))
    
#filtered_Mitigation_nat1 <- subset(filtered_Mitigation_nat, select = -c(ID, Applies.To, Exceptions, Implementing.City, Implementing.State.Province, Quantity, Source, Target.city, Target.country, Target.region, Target.state))
    
filtered_Mitigation_gov1 <- subset(filtered_Mitigation_gov, select = -c(ID, ADMIN_LEVEL_NAME, PCODE, REGION, MEASURE, COMMENTS, NON_COMPLIANCE, SOURCE, SOURCE_TYPE, LINK, Alternative.source))


```




##filter by date
```{r}


```



#exporting Cleaned data
```{r}
write.csv(filtered_Mitigation_gov1, file = "filtered_Mitigation_gov1")

```



#############################################################################################################
#############################################################################################################

#covid measures
```{r}
covid.19.all <- read.csv("C:/Users/micha/Desktop/Semester8/Introduction to Data Sience/Project_Group/Neuer Ordner/Data/covid-19-all.csv")

#View(covid.19.all)
#head(covid.19.all)
#tail(covid.19.all)
summary(covid.19.all)
str(covid.19.all)
#count(covid.19.all$Country.Region,"Switzerland")

#colums cleaning
new_covid19_all <- subset(covid.19.all, select =-c(Latitude:Longitude))
summary(new_covid19_all)

#row cleaning
#emty data frames
new_Switzerland_covid19_all <- data.frame()
new_Italy_covid19_all <- data.frame()
new_Germany_covid19_all <- data.frame()
new_UnitedKingdom_covid19_all <- data.frame()
new1_covid19_all <- data.frame()

#filter for countries 
library(dplyr)
vec=c("Switzerland","Italy","Germany", "United Kingdom")
#vec=c("United Kingdom", "Germany")
new1_covid19_all <- filter(new_covid19_all, Country.Region %in% vec)

new_Switzerland_covid19_all <- filter(new_covid19_all, Country.Region %in% "Switzerland")

new_Italy_covid19_all <- filter(new_covid19_all, Country.Region %in% "Italy")

new_UnitedKingdom_covid19_all <-filter(new_covid19_all, Country.Region %in% "United Kingdom")

new_Germany_covid19_all <- filter(new_covid19_all, Country.Region %in% "Germany")

#dataset for Switzerland, Germany and Italy
```

##make sure dates are safed as dates
```{r}
new_Switzerland_covid19_all$Date <- as.Date(new_Switzerland_covid19_all$Date, format = "%Y-%m-%d")
new_Italy_covid19_all$Date <- as.Date(new_Italy_covid19_all$Date, format = "%Y-%m-%d")
new_UnitedKingdom_covid19_all$Date <-as.Date(new_UnitedKingdom_covid19_all$Date, format = "%Y-%m-%d")
new_Germany_covid19_all$Date <-as.Date(new_Germany_covid19_all$Date, format = "%Y-%m-%d")
new1_covid19_all$Date <- as.Date(new1_covid19_all$Date, format = "%Y-%m-%d")
```

#cleaning data _measurements: april(29.3.2020-30.4.2020) + august 2020-07-24 23:47:08 (24.7.2020-30.8.2020)
```{r}
measurementrs_Covid_all <- data.frame()
measurementrs_Covid_Switzerland <-data.frame()
measurementrs_Covid_Italy <- data.frame()
measurementrs_Covid_UnitedKingdom <- data.frame()
measurementrs_Covid_Germany <- data.frame()

##Covid19 Measurements_all
cov1_all_april <- filter(new1_covid19_all, between(Date, as.Date("2020-03-29"), as.Date("2020-04-30")))
cov1_all_august <- filter(new1_covid19_all, between(Date, as.Date("2020-07-24"), as.Date("2020-08-30")))
measurementrs_Covid_all <- full_join(cov1_all_april, cov1_all_august)
View(measurementrs_Covid_all)

###covid19 measurements_Switzerland
cov1_Swiss_april <- filter(new_Switzerland_covid19_all, between(Date, as.Date("2020-03-29"), as.Date("2020-04-30")))
cov1_Swiss_august <- filter(new_Switzerland_covid19_all, between(Date, as.Date("2020-07-24"), as.Date("2020-08-30")))
measurementrs_Covid_Switzerland <- full_join(cov1_Swiss_april, cov1_Swiss_august)
View(measurementrs_Covid_Switzerland)

###covid 19 measurements_Italy
cov1_Italy_april <- filter(new_Italy_covid19_all, between(Date, as.Date("2020-03-29"), as.Date("2020-04-30")))
cov1_Italy_august <- filter(new_Italy_covid19_all, between(Date, as.Date("2020-07-24"), as.Date("2020-08-30")))
measurementrs_Covid_Italy <- full_join(cov1_Italy_april, cov1_Italy_august)
View(measurementrs_Covid_Italy)

##covid 19 measurements_united kingdom
cov1_UK_april <- filter(new_UnitedKingdom_covid19_all, between(Date, as.Date("2020-03-29"), as.Date("2020-04-30")))
cov1_UK_august <- filter(new_UnitedKingdom_covid19_all, between(Date, as.Date("2020-07-24"), as.Date("2020-08-30")))
measurementrs_Covid_UnitedKingdom <- full_join(cov1_UK_april, cov1_UK_august)
View(measurementrs_Covid_UnitedKingdom)

##covid 19 measurements_germany
cov1_germ_april <- filter(new_Germany_covid19_all, between(Date, as.Date("2020-03-29"), as.Date("2020-04-30")))
cov1_germ_august <- filter(new_Germany_covid19_all, between(Date, as.Date("2020-07-24"), as.Date("2020-08-30")))
measurementrs_Covid_Germany <- full_join(cov1_germ_april, cov1_germ_august)
View(measurementrs_Covid_Germany)



#date <- c("2020-03-29":"2020-04-30", "2020-07-24":"2020-08-30")
#View(date)
#measurementrs_Covid_all <- filter(new1_covid19_all, Date %in% date)
```



#exporting Cleaned data
```{r}
write.csv(measurementrs_Covid_all, file = "measurementrs_Covid_all")
write.csv(measurementrs_Covid_Switzerland, file = "measurementrs_Covid_Switzerland")
write.csv(measurementrs_Covid_Italy, file = "measurementrs_Covid_Italy")
write.csv(measurementrs_Covid_Germany, file = "measurementrs_Covid_Germany")
write.csv(measurementrs_Covid_UnitedKingdom, file = "measurementrs_Covid_UnitedKingdom")

```






